---
title: 'Chapter 3: Compare means'
author: "Khia A. Johnson"
date: "7/23/2021"
output: html_document
---

This file accompanies the comparison of means using Cohen's d in Chapter 3, and also does all the data viz for that section.

---

# Set up

Imports
```{r}
library(data.table)
library(tidyverse)
library(ggthemes)
library(lsr)
library(broom)
library(ggdist)
```

Set visualization defaults
```{r}
theme_set(theme_clean() + theme(
      legend.title = element_blank(),
      legend.text = element_text(size=6),
      legend.position = 'bottom',
      axis.title.x = element_text(size = 10, face='bold'),
      axis.title.y = element_text(size = 10, face='bold'),
      strip.text = element_text(size=10, face='bold'),
      plot.background = element_blank()
    ))

options(scipen = 999)
```

Load the processed data frame generated in `3a-wrangle.rmd`
```{r}
df <- fread('../data/spice_voicesauce_processed.csv', sep = ',') 
```

Compute the correlation for F4 and FD
```{r}
fd <- df %>%
    select(F1:F4) %>%
    mutate(FD = ((F4-F3) + (F3-F2) + (F2-F1) )/3)

cor.test(x = fd$F4, y= fd$FD, method = 'pearson')
```

# Visualize distributions

Pivot longer to prep for getting t-test and cohen's d results // data viz
```{r}
long_df <- df %>%
  select(Talker, Language, F0:SHR_sd) %>%
  pivot_longer(F0:SHR_sd, names_to = 'Variable', values_to = 'Value')

long_df 
```

```{r}
long_df_main <- long_df %>%
    mutate(
        Variable = str_replace(Variable, 'H1H2c', 'H1*-H2*'),
        Variable = str_replace(Variable, 'H2H4c', 'H2*-H4*'),
        Variable = str_replace(Variable, 'H42Kc', 'H4*-H2kHz*'),
        Variable = str_replace(Variable, 'H2KH5Kc', 'H2kHz*-H5kHz')
    ) %>%
    filter(str_detect(Variable, 'sd', negate=TRUE)) %>%
    mutate(Variable = as_factor(Variable))
```


```{r fig.width=5, fig.height=6.5}
long_df_main %>%
  select(Talker, Language, Variable, Value) %>%
  ggplot(aes(x=Value, color=Language)) +
    geom_density() +
    scale_color_viridis_d(option='magma', begin = 0.75, end = 0)+
    facet_wrap(~Variable, scales = 'free', ncol = 3) +
    ylab('Density') +
    xlab('')

#ggsave('../../../text/figures/ch3_allmeasuresdensity_5in.png', width = 5, height = 6.5)

```

Inspect highly skewed SHR
```{r fig.width=5, fig.height=6.5}
long_df_main %>%
  select(Talker, Language, Variable, Value) %>%
  filter(Variable=='SHR') %>%
  ggplot(aes(x=Value, color=Language)) +
    geom_density() +
    scale_color_viridis_d(option='magma', begin = 0.75, end = 0)+
    facet_wrap(~Talker, ncol = 7) +
    ylab('Density') +
    xlab('')
```

Inspect skewed Energy
```{r fig.width=5, fig.height=6.5}
long_df_main %>%
  select(Talker, Language, Variable, Value) %>%
  filter(Variable=='Energy') %>%
  ggplot(aes(x=Value, color=Language)) +
    geom_density() +
    scale_color_viridis_d(option='magma', begin = 0.75, end = 0)+
    facet_wrap(~Talker, ncol = 7) +
    ylab('Density') +
    xlab('')
```


# Cohen's d

Calculate cohen's d and means, then combine and save results

```{r}
cohens_d_results <- long_df %>%
    group_by(Talker, Variable) %>%
    summarize(cohen_d = cohensD(Value ~ Language, method='pooled'))

means <- long_df %>%
    group_by(Talker, Variable, Language) %>%
    summarise(mv = mean(Value)) %>%
    pivot_wider(names_from = Language, values_from = mv)

cohens_d_results <- cohens_d_results %>%
    left_join(means, by=c('Talker','Variable')) %>%
    select(Talker, Variable, CohensD=cohen_d, C_mean=Cantonese, E_mean=English) 

fwrite(cohens_d_results, '../data/spice_voicesauce_cohensd.csv')
cohens_d_results <- fread('../data/spice_voicesauce_cohensd.csv')
```

Add qualitative effect size groupings and inspect
```{r fig.width=6, fig.height=12}
cohens_d_results <- cohens_d_results %>%
  mutate(Difference = C_mean-E_mean,
         CohensD_Group = if_else(CohensD<0.2, 'Trivial', 
                          if_else(CohensD<0.5, 'Small', 
                                  if_else(CohensD<0.8, 'Medium', 'Large')))
         ) 

cohens_d_results
```

Count the number of non-trivial comparisons by talker and output a histogram
```{r}
cohens_d_results %>%
  filter(CohensD_Group != 'Trivial') %>%
  group_by(Talker) %>%
  summarise(n=n()) %>%
  ggplot(aes(x=n)) +
  geom_histogram(binwidth = 1, color='white', fill='black') +
  scale_x_continuous(limits = c(0,13), breaks = c(0,1,2,3,4,5,6,7,8,9,10,11,12) ) +
  ylab('Count') +
  xlab('Number of non-trivial comparisons')

#ggsave('../../../text/figures/ch3_nontrivial_counts_by_talker_5in.png', width = 5, height = 2.5)
```

Counts by Cohen's d brackets for table
```{r}
cohens_d_results %>%
  mutate(positive = Difference>0) %>%
  group_by(Variable, positive, CohensD_Group) %>%
  summarise(n = n()) %>%
  pivot_wider(names_from = c('positive','CohensD_Group'), values_from = n) %>%
  select(Variable, FALSE_Small, FALSE_Medium, TRUE_Small, TRUE_Medium)
```

## Data visualization

Prep data frame for data visualization
```{r}
long_pointinterval <- long_df %>%
    group_by(Talker, Language, Variable) %>%
    summarise(m = mean(Value), 
              q025 = quantile(Value, 0.025), 
              q25 = quantile(Value, 0.25), 
              q75= quantile(Value, 0.75), 
              q975 = quantile(Value, 0.975)) %>%
    left_join(cohens_d_results, by=c('Talker','Variable')) %>%
    mutate(
        CohensD_Group = as_factor(CohensD_Group),
        Variable = str_replace(Variable, '_sd', ' s.d.'),
        Variable = str_replace(Variable, 'H1H2c', 'H1*-H2*'),
        Variable = str_replace(Variable, 'H2H4c', 'H2*-H4*'),
        Variable = str_replace(Variable, 'H42Kc', 'H4*-H2kHz*'),
        Variable = str_replace(Variable, 'H2KH5Kc', 'H2kHz*-H5kHz'),
        Variable = as_factor(Variable))

f <- c("F0", "F0 s.d.", "F1", "F1 s.d.", "F2", "F2 s.d.", "F3", "F3 s.d.",
              "F4", "F4 s.d.","H1*-H2*", "H1*-H2* s.d.","H2*-H4*", "H2*-H4* s.d.",
              "H4*-H2kHz*", "H4*-H2kHz* s.d.","H2kHz*-H5kHz", "H2kHz*-H5kHz s.d.",
              "CPP", "CPP s.d.", "Energy", "Energy s.d.","SHR", "SHR s.d.")
# 
fwrite(long_pointinterval, '../data/spice_long_pointinterval.csv')
```

Plot cohen's d by difference across means, part 1
```{r fig.height=6.5, fig.width=5}
long_pointinterval %>%
  mutate(Variable = factor(Variable, levels = f), CohensD_Group = factor(CohensD_Group, levels = c('Trivial','Small','Medium'))) %>%
  filter(str_detect(Variable, 's.d.', negate=TRUE)) %>%
  ggplot(aes(x=CohensD, y=Difference, color=CohensD_Group, shape=CohensD_Group)) + 
  geom_point() +
  geom_hline(aes(yintercept = 0))+
  scale_color_viridis_d(option='magma', begin = 0.9, end=0) +
  facet_wrap(~Variable, scales = 'free', ncol=3)

ggsave('../../../text/figures/ch3_cohend_part1_5in.png', width = 5, height = 6.5)
```

Plot cohen's d by difference across means, part 2
```{r fig.height=6.5, fig.width=5}
long_pointinterval %>%
  mutate(Variable = factor(Variable, levels = f), CohensD_Group = factor(CohensD_Group, levels = c('Trivial','Small','Medium'))) %>%
  filter(str_detect(Variable, 's.d.', negate=FALSE)) %>%
  ggplot(aes(x=CohensD, y=Difference, color=CohensD_Group, shape=CohensD_Group)) + 
  geom_point() +
  geom_hline(aes(yintercept = 0))+
  scale_color_viridis_d(option='magma', begin = 0.9, end=0) +
  facet_wrap(~Variable, scales = 'free', ncol=3)

ggsave('../../../text/figures/ch3_cohend_part2_5in.png', width = 5, height = 6.5)
```

# Inspect F2

```{r}
f2d <- cohens_d_results %>%
    filter(Variable=='F2') %>%
    select(Talker, CohensD, Difference)
```

```{r fig.width=10, fig.height=6.5}
df %>%
  select(Talker, Language, F2) %>%
  left_join(f2d, by='Talker') %>%
  mutate(Difference = Difference>0) %>%
  ggplot(aes(x=F2, color=Language)) +
    geom_density() +
    scale_color_viridis_d(option='magma', begin = 0.75, end = 0)+
    facet_wrap(~Talker, ncol = 8) +
    ylab('Density') +
    xlab('')
```
# Inspect F0
```{r}
f0d <- cohens_d_results %>%
    filter(Variable=='F0') %>%
    select(Talker, CohensD, Difference)
```


```{r}
f0d %>%
    mutate(nontrivial = abs(CohensD)>=0.2) %>%
    filter(nontrivial==TRUE)
```



```{r fig.width=6, fig.height=3}
df %>%
  select(Talker, Language, F0) %>%
  left_join(f0d, by='Talker') %>%
  mutate(Difference = Difference>0) %>%
    mutate(nontrivial = abs(CohensD)>=0.2) %>%
    filter(nontrivial==TRUE) %>%
  ggplot(aes(x=F0, color=Language, linetype=Language)) +
    geom_density(size=1) +
    scale_color_viridis_d(option='magma', begin = 0.75, end = 0)+
    facet_wrap(~Talker, ncol = 6) +
    scale_x_continuous(limits = c(70,325)) +
    ylab('Density') +
    xlab('')
    
ggsave('../../../text/figures/ch3_f0_means_10in.png', width = 10, height = 5.5)
```

---