---
title: "4c-cca"
author: "Khia A. Johnson"
date: "8/17/2021"
output: html_document
---

This file accompanies the canonical correlation analysis in Chapter 3.

---

# Setup

Imports
```{r}
library(tidyverse)
library(ggthemes)
library(ggrepel)
library(parameters)
library(candisc)
```

Set visualization defaults
```{r}
theme_set(theme_clean() + theme(
      legend.title = element_blank(),
      legend.text = element_text(size=6),
      legend.position = 'bottom',
      axis.title.x = element_text(size = 10, face='bold'),
      axis.title.y = element_text(size = 10, face='bold'),
      strip.text = element_text(size=10, face='bold'),
      plot.background = element_blank()
    ))

options(scipen = 999)
```

Load data
```{r}
results <- read_csv('../data/pca_results_truncated.csv')
files <- unique(results$File)
```

# Canonical Correlation Analysis

## Run CCA

Function to get a tibble of a single PCA's loadings matrix given the filename
```{r}
get_pca_loadings_tibble <- function(f) {
  m <- results %>%
    filter(File==f) %>%
    select(Variable, Var_Order, Loading) %>%
    arrange(Var_Order, Variable) %>%
    pivot_wider(names_from = Var_Order, values_from = Loading)
  return(m[,2:ncol(m)]) 
}
```

Run CCA and get redundancy indices for all unique combinations, and save it
```{r}
all_pairs <- t(combn(files,2))

xs <- list()
ys <- list()
px <- list()
py <- list()

for (i in 1:nrow(all_pairs)) {

  x_file <- all_pairs[[i, 1]]
  y_file <- all_pairs[[i, 2]]
  this_CCA <- cancor(
    x=get_pca_loadings_tibble(x_file),
    y=get_pca_loadings_tibble(y_file))
  this_red <- redundancy(this_CCA)

  px[[i]] <- x_file
  py[[i]] <- y_file
  xs[[i]] <- this_red$X.redun
  ys[[i]] <- this_red$Y.redun
  }

all_redundancy <- as_tibble(cbind(px,py,xs,ys)) %>% 
  unnest(c(px,py,xs,ys))

rm(px, py, xs, ys, this_CCA, this_red, this_talker, i, all_pairs, x_file, y_file)

all_redundancy %>%
  write_csv('../data/truncated_all_redundancy_indices.csv')
```

Load (if already run)
```{r}
all_redundancy <- read_csv('../data/truncated_all_redundancy_indices.csv')
```

Modify redundancy dataframe before plotting, and inspect it
```{r}
all_redundancy <- all_redundancy %>%
  separate(px, into = c('x_talker','x_language')) %>%
  separate(py, into = c('y_talker','y_language')) %>%
  mutate(x_language = str_replace(x_language, '[0-9]', ''),
         y_language = str_replace(y_language, '[0-9]', ''),
         same_language = x_language == y_language,
         same_talker = x_talker == y_talker)
  
head(all_redundancy)
```  
 
## Analyze CCA
 
Visualize x and y redundancy by type of comparison
```{r fig.width=5, fig.height=5}
all_redundancy %>%
    mutate(grp = paste0(same_talker,same_language),
           grp = str_replace(grp, 'TRUEFALSE', 'Same talker, different language'),
           grp = str_replace(grp, 'FALSETRUE', 'Different talker, same language'),
           grp = str_replace(grp, 'FALSEFALSE', 'Different talker, different language')
    ) %>%
  ggplot(aes(x=xs, y=ys, color=grp,fill=grp, shape=grp)) +
        geom_point(size=4) + 
        scale_color_viridis_d(alpha=0.6,option = 'magma', begin=0.85, end=0) +
        scale_fill_viridis_d(alpha=0.9,option = 'magma', begin=0.85, end=0) +
        scale_shape_manual(values = c(3,4,21)) +
        xlab('Redundancy (x)') +
        ylab('Redundancy (y)') +
        theme_clean() +
        theme(legend.title = element_blank(), 
              legend.position = 'bottom') +
     guides(color=guide_legend(nrow=3))

# ggsave('../../../text/figures/ch3_redundancy.png', width = 5, height = 5)

```
 
Numerical summary of groups
```{r}
all_redundancy %>%
    mutate(grp = paste0(same_talker,same_language),
           grp = str_replace(grp, 'TRUEFALSE', 'Same talker, different language'),
           grp = str_replace(grp, 'FALSETRUE', 'Different talker, same language'),
           grp = str_replace(grp, 'FALSEFALSE', 'Different talker, different language')
    ) %>%
select(grp, xs, ys) %>%
    pivot_longer(xs:ys) %>%
    group_by(grp) %>%
    summarise(min(value), max(value), mean(value), median(value), sd(value)) 
```

Pivot and do a t-test for same/diff talker
```{r}
red_pivoted <- all_redundancy %>%
    select(same_talker, xs, ys) %>%
    pivot_longer(xs:ys)

t.test(value ~ same_talker, data = red_pivoted)
```

Pivot and do a t-test for language across talker
```{r}
red_pivoted2 <- all_redundancy %>%
    filter(same_talker==FALSE) %>%
    select(same_language, xs, ys) %>%
    pivot_longer(xs:ys)

t.test(value ~ same_language, data = red_pivoted2)
```

## Plot redundancy by test diff

Load data frame from `3b-means.rmd` file
```{r}
long_pointinterval <- read_csv('../data/spice_long_pointinterval.csv')
```

Join redundancy and cohen's d info
```{r fig.width=5, fig.height=6}
redbydiff <- all_redundancy %>%
  filter(same_talker==TRUE, same_language==FALSE) %>%
  mutate(average_redundancy = (xs+ys)/2) %>%
  select(Talker=x_talker, average_redundancy) %>%
  left_join(long_pointinterval, by='Talker') %>%
  mutate(
    CohensD = as_factor(CohensD),
    abs_difference = abs(Difference),
    average_redundancy = round(average_redundancy,2),
    Variable = str_replace(Variable, '_sd', ' s.d.'),
    Variable = str_replace(Variable, 'H1H2c', 'H1*-H2*'),
    Variable = str_replace(Variable, 'H2H4c', 'H2*-H4*'),
    Variable = str_replace(Variable, 'H42Kc', 'H4*-H2kHz*'),
    Variable = str_replace(Variable, 'H2KH5Kc', 'H2kHz*-H5kHz'),
    Variable = as_factor(Variable)
  ) %>%
  filter(str_detect(Variable, 's.d.', negate=TRUE)) 
```

```{r}
redbydiff
```

```{r}
f <- c("F0", "F0 s.d.", "F1", "F1 s.d.", "F2", "F2 s.d.", "F3", "F3 s.d.",
              "F4", "F4 s.d.","H1*-H2*", "H1*-H2* s.d.","H2*-H4*", "H2*-H4* s.d.",
              "H4*-H2kHz*", "H4*-H2kHz* s.d.","H2kHz*-H5kHz", "H2kHz*-H5kHz s.d.",
              "CPP", "CPP s.d.", "Energy", "Energy s.d.","SHR", "SHR s.d.")
```


Plot the comparison
```{r fig.width=5, fig.height=6}
redbydiff %>%
  mutate(Variable = factor(Variable, levels = f), CohensD_Group = factor(CohensD_Group, levels = c('Trivial','Small','Medium'))) %>%
  ggplot(aes(x=abs_difference, y=average_redundancy)) + 
  geom_point(aes(color=CohensD_Group, shape=CohensD_Group)) +
  stat_smooth(method = 'lm', color='black', fill='black', alpha=0.3) +
  scale_y_continuous(limits = c(0.85, 0.97),breaks = c(0.85, 0.9, 0.95)) +
  scale_color_viridis_d(option='magma', begin = 0.9, end=0) +
  facet_wrap(~Variable, scales = 'free_x', ncol=3) +
  ylab('Redundancy (Average)') + 
  xlab('Absolute Mean Difference') +
    theme(legend.position = 'bottom')

ggsave('../../../text/figures/ch3_redundancybyttestdiff_5in.png', width = 5, height = 6)
```

