---
title: "Chapter 4: Pairwise correlations"
author: "Khia A Johnson"
date: "4/1/2021"
output: html_document
---

This file accompanies the pairwise correlations part of Chapter 4
---

# Setup
Imports
```{r}
library(tidyverse)
library(ggthemes)
library(gridExtra)
library(psych)
library(kableExtra)
library(ggExtra)
library(scales)
```

Viz defaults
```{r}
options(scipen = 999)

theme_set(
  theme_clean() + 
  theme(
      legend.title = element_blank(),
      legend.text = element_text(size=8),
      legend.position = 'bottom',
      axis.title.x = element_text(size = 8, face='bold'),
      axis.title.y = element_text(size = 8, face='bold'),
      axis.text.x = element_text(size = 8),
      axis.text.y = element_text(size = 8),
      strip.text = element_text(size = 8, face='bold'),
      )
  )
```

Get color hexes from theme for manual setting later
```{r}
pal <- viridis_pal(option='magma', begin=0, end=0.85)
show_col(pal(4))
pal(4)
```

Load data
```{r}
df <- read_csv('../data/ptk_clean_20210602.csv') # this version used in Interspeech paper
```

Get means
```{r}
means <- df %>%
  group_by(talker, language, segment) %>%
  summarise(mean_vot_ms = mean(vot_ms)) %>%
  ungroup() %>%
  pivot_wider(names_from = 'segment', values_from = 'mean_vot_ms') 
```

# Correlations

## Raw VOT 

### Calculate

Prep wide format data frame
```{r}
dfsummary <- df %>%
  group_by(talker, language, segment) %>%
  summarise(mvot = mean(vot_ms)) %>%
  ungroup() %>%
  pivot_wider(names_from = c('language','segment'), values_from = 'mvot') %>%
  select(
    `Cantonese P`=Cantonese_p, 
    `Cantonese T`=Cantonese_t, 
    `Cantonese K`=Cantonese_k, 
    `English P`=English_p, 
    `English T`=English_t, 
    `English K`=English_k)
```

Prep wide summary data frame and run correlation
```{r}
dfout <- corr.test(dfsummary, method='pearson', adjust = 'holm')

dfout$ci %>%
  rownames_to_column(var = 'comparison') %>%
  left_join(dfout$ci2, by=c('r','p')) %>%
  select(comparison, r, p.adj) %>%
  separate(comparison, into = c('a', 'b'), sep='-') %>%
  mutate(r = round(r, 2), p.adj = round(p.adj, 4)) 
```

Run correlations with adjusted p-values and inspect output
```{r}
output <- corr.test(dfsummary, method='pearson', adjust = 'holm')
output
```

### Visualize

```{r}
jointplot <- function(dataframe, col1, name1, col2, name2, color='black') {
  p <- ggplot(dataframe, aes(x=col1, y=col2)) +
    geom_point(color=color) +
    geom_smooth(method='lm', formula = 'y ~ x', color=color, fill=color) +
    scale_x_continuous(limits = c(10,90)) + 
    scale_y_continuous(limits = c(10,90)) +
    xlab(paste(name1, '(Mean VOT)', sep = ' ')) +
    ylab(paste(name2,'(Mean VOT)', sep = ' ')) +
    theme(
      axis.title.x = element_text(size = 7, face='bold'),
      axis.title.y = element_text(size = 7, face='bold'),
    )
  
  x <- cor.test(x=col1, y=col2)
  x <- paste("r =",as.character(round(x$estimate,2)))
  
  p <- p + annotate("text", x = 60, y = 15, label = x)

  p <- ggMarginal(p, type="histogram", fill=color, color='white') 

  return(p)
}
```

```{r}
cpt <- jointplot(dfsummary, dfsummary$`Cantonese P`, 'Cantonese /p/', dfsummary$`Cantonese T`, 'Cantonese /t/', color='#000004FF')
cpk <- jointplot(dfsummary, dfsummary$`Cantonese P`, 'Cantonese /p/', dfsummary$`Cantonese K`, 'Cantonese /k/', color='#000004FF')
ctk <- jointplot(dfsummary, dfsummary$`Cantonese T`, 'Cantonese /t/', dfsummary$`Cantonese K`, 'Cantonese /K/', color='#000004FF')
ept <- jointplot(dfsummary, dfsummary$`English P`, 'English /p/', dfsummary$`English T`, 'English /t/', color='#5D177FFF')
epk <- jointplot(dfsummary, dfsummary$`English P`, 'English /p/', dfsummary$`English K`, 'English /k/', color='#5D177FFF')
etk <- jointplot(dfsummary, dfsummary$`English T`, 'English /t/', dfsummary$`English K`, 'English /K/', color='#5D177FFF')
cep <- jointplot(dfsummary, dfsummary$`Cantonese P`, 'Cantonese /p/', dfsummary$`English P`, 'English /p/', color='#D1426FFF')
cet <- jointplot(dfsummary, dfsummary$`Cantonese T`, 'Cantonese /t/', dfsummary$`English T`, 'English /t/', color='#D1426FFF')
cek <- jointplot(dfsummary, dfsummary$`Cantonese K`, 'Cantonese /k/', dfsummary$`English K`, 'English /K/', color='#D1426FFF')
cpet <- jointplot(dfsummary, dfsummary$`Cantonese P`, 'Cantonese /p/', dfsummary$`English T`, 'English /t/', color='#FEB77EFF')
cpek <- jointplot(dfsummary, dfsummary$`Cantonese P`, 'Cantonese /p/', dfsummary$`English K`, 'English /k/', color='#FEB77EFF')
ctep <- jointplot(dfsummary, dfsummary$`Cantonese T`, 'Cantonese /t/', dfsummary$`English P`, 'English /p/', color='#FEB77EFF')
ctek <- jointplot(dfsummary, dfsummary$`Cantonese T`, 'Cantonese /t/', dfsummary$`English K`, 'English /k/', color='#FEB77EFF')
ckep <- jointplot(dfsummary, dfsummary$`Cantonese K`, 'Cantonese /k/', dfsummary$`English P`, 'English /p/', color='#FEB77EFF')
cket <- jointplot(dfsummary, dfsummary$`Cantonese K`, 'Cantonese /k/', dfsummary$`English T`, 'English /t/', color='#FEB77EFF')
```

```{r}
p <- grid.arrange(cpt, cpk, ctk, ept, epk, etk, ncol=3)
# ggsave('../../text/figures/ch4_correlations1_5in.png', plot=p, height = 4, width = 6, units='in')
```

```{r}
p <- grid.arrange(cep, cet, cek, cpet, cpek, ctep, ctek, ckep, cket, ncol=3)
# ggsave('../../text/figures/ch4_correlations2_5in.png',plot=p, height = 6, width = 6, units='in')

```


## Residual VOT (accounting for rate)

Compute residual VOT via simple VOT ~ rate regression, and add to data frame
```{r}
rdf <- df
resmodel <- lm(vot_ms ~ avg_phone_dur_ms, data = rdf)
rdf$residual_vot_ms <- resid(resmodel)
```

Prep wide summary data frame and run correlation
```{r}
rdfsummary <- rdf %>%
  group_by(talker, language, segment) %>%
  summarise(mvot = mean(residual_vot_ms)) %>%
  ungroup() %>%
  pivot_wider(names_from = c('language','segment'), values_from = 'mvot') %>%
  select(
    `Cantonese P`=Cantonese_p, 
    `Cantonese T`=Cantonese_t, 
    `Cantonese K`=Cantonese_k, 
    `English P`=English_p, 
    `English T`=English_t, 
    `English K`=English_k)
 
rdfout <- corr.test(rdfsummary, method='pearson', adjust = 'holm')
```

Print the table output for the residual VOT comparisons
```{r}
rdfout$ci %>%
  rownames_to_column(var = 'comparison') %>%
  left_join(rdfout$ci2, by=c('r','p')) %>%
  select(comparison, r, p.adj) %>%
  separate(comparison, into = c('a', 'b'), sep='-') %>%
  mutate(r = round(r, 2), p.adj = round(p.adj, 4)) 
```

---