---
title: "Chapter 2: The SpiCE Corpus"
author: "Khia A. Johnson"
date: "5/21/2021"
output: html_document
---

This file accompanies chapter 2 of my dissertation. It includes descriptive statistics and data visualization summarizing the participants and contents of the corpus in broad strokes.

---

# Setup

Imports
```{r}
library(tidyverse)
library(ggthemes)
library(GGally)
library(ggrepel)
library(DBI)
library(viridis)
```

Set visualization defaults
```{r}
theme_set(theme_clean() + theme(
      legend.title = element_text(face = 'bold', size=10),
      legend.text = element_text(size=10),
      legend.position = 'bottom',
      axis.title.x = element_text(size = 10, face='bold'),
      axis.title.y = element_text(size = 10, face='bold'),
      strip.text = element_text(size=10, face='bold'),
      plot.background = element_blank()
    ))
```

Connect to the SpiCE SQLite database created in `../1-general/setup-spice-db.py` file, and print out the names of the tables.
```{r}
db <- dbConnect(RSQLite::SQLite(), "../1-general/spice.db")
dbListTables(db)
```

Load the words, tasks, and phones tables
```{r}
words <- dbReadTable(db, "words")
tasks <- dbReadTable(db, "tasks")
phones <- dbReadTable(db, "phones")
```

Query the full words table merged with tasks by timestamps with SQL query and grab variable info from file names.
```{r}
results <- dbSendQuery(
    db,
    "SELECT word,
	       word_onset,
	       word_offset,
	       task,
	       task_onset,
	       task_offset,
	       words.file
	FROM words
	LEFT JOIN tasks ON (tasks.file = words.file
	                    AND word_onset BETWEEN task_onset AND task_offset)"
    )

words <- fetch(results)

words <- words %>%
    separate(file, into = c('talker','language','order','date'), remove = FALSE)
```

Take a peek at a common Canadian/American spelling distinction... unfortunately it's not 100% consistent.
```{r}
words %>%
    group_by(word) %>%
    summarise(n()) %>%
    filter(str_detect(word,'our'))
```

# Corpus Summary

## Counting words

Query a table with both utterance and task information, and inspect it.
```{r}
results <- dbSendQuery(
    db,
    "SELECT utterance,
	       utterance_onset,
	       utterance_offset,
	       task,
	       task_onset,
	       task_offset,
	       utterances.file
	FROM utterances
	LEFT JOIN tasks ON (tasks.file = utterances.file
	                    AND utterance_onset BETWEEN task_onset AND task_offset)"
    )
utts <- fetch(results)

utts <- utts %>%
    separate(file, into = c('talker','language','order','date'), remove = FALSE)

utts
```

Create a table of counting words by talker according to type of word: English, Cantonese, Other language, Fragment, or Unintelligible.
```{r}
word_counts <- utts %>%
    filter(task == 'interview') %>%
    select(utt=utterance, talker, interview=language) %>%
    mutate(
        utt = str_replace_all(utt, '\\s*[@＠]\\s*', '@'),
        utt = str_replace_all(utt, '[】 【]', ' '),
        utt = str_replace_all(utt, '\\s+', ' '),
        utt = str_replace(utt, '^\\s+', ''),
        utt = str_replace(utt, '\\s+$', '')
        ) %>%
    separate_rows(utt, sep = '\\s+') %>%
    group_by(utt, talker, interview) %>%
    summarise(n=n()) %>%
    mutate(
        utt = str_replace_all(utt, '[\\?}@＠]$', ''),
        type = if_else(str_starts(utt, '&'), 'Fragment', NULL),
        type = if_else(str_detect(utt, '@'), 'Other Language', type), 
        type = if_else(str_starts(utt, '[a-z]') & str_ends(utt, '[0-9]'), 'Cantonese', type), 
        type = if_else(str_detect(utt, "^[a-z'-]+$"), 'English', type), 
        type = if_else(utt == 'xxx', 'Unintelligible', type),
        type = if_else(is.na(type), 'Cantonese', type)
    ) %>%
    filter(utt != '', str_starts(utt, '@', negate = TRUE)) %>%
    select(word = utt, talker, interview, n, type)
```

Generate a summary table for the word counts according to tyep of word for use in plotting.
```{r}
word_count_summary <- word_counts %>%
    group_by(interview, talker, type) %>%
    summarise(n=sum(n)) %>%
    ungroup()
```

Figure showing proportion of word types by talker across Cantonese interviews.
```{r, fig.width=5, fig.height=7}
word_count_summary %>%
    filter(interview=='Cantonese') %>%
    ggplot(aes(fill=type, x=talker, y=n)) + 
    geom_col() + 
    scale_fill_viridis_d(option='magma', begin = 0.9, end = 0,
                         limits = c('Cantonese','English','Other Language', 'Fragment', 'Unintelligible')) + 
    xlab('Talker') +
    ylab('Number of words') +
    theme(axis.text.x = element_text(angle = 90),
          legend.position="bottom",
          legend.title = element_blank(),
          legend.text = element_text(size=6)) 

# ggsave('../../text/figures/ch2_cantonesetypecounts_5in.png', width = 5, height = 6.5)
```

Figure showing proportion of word types by talker across English interviews.
```{r, fig.width=5, fig.height=7}
word_count_summary %>%
    filter(interview=='English') %>%
    ggplot(aes(fill=type, x=talker, y=n)) + 
    geom_col() + 
    scale_fill_viridis_d(option='magma', begin = 0.9, end = 0,
                         limits = c('Cantonese','English','Other Language', 'Fragment', 'Unintelligible')) + 
    xlab('Talker') +
    ylab('Number of words') +
    theme(axis.text.x = element_text(angle = 90),
          legend.position="bottom",
          legend.title = element_blank(),
          legend.text = element_text(size=6)) 

# ggsave('../../text/figures/ch2_englishtypecounts_5in.png', width = 5, height = 6.5)
```

Unique number of words in each language
```{r}
word_counts %>%
    ungroup() %>%
    select(word,interview) %>%
    unique() %>%
    group_by(interview) %>%
    summarise(n()) 
```

Token and type count information summarizing across talkers.
```{r}
word_counts %>%
    group_by(interview, talker) %>%
    summarise(tokens = sum(n), types = n()) %>%
    group_by(interview) %>%
    summarise(sum(tokens),mean(tokens),sd(tokens), min(tokens),max(tokens), 
             mean(types),sd(types), min(types),max(types))
```

Word frequency

Plot log word frequency for both languages, excluding instances of <unk>
```{r fig.width=6, fig.height=3}
words %>%
    filter(word != '<unk>') %>%
    group_by(language, word) %>%
    summarise(frequency = n()) %>%
    ungroup() %>%
    mutate(log_frequency = log(frequency)) %>%
    ggplot(aes(x=log_frequency)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_wrap(~language) +
        xlab('Log frequency') +
        ylab('Count')

# ggsave('../../text/figures/ch2_wordfrequency_5in.png', width = 5, height = 3)
```

## Language switching

English interviews with fewer than 10 Cantonese words
```{r}
word_counts %>%
    group_by(talker, interview, type) %>%
    summarise(n=sum(n)) %>%
    pivot_wider(names_from = 'type', values_from = 'n') %>%
    filter(Cantonese<10)
```

English and Cantonese token frequency in the Cantonese sessions
```{r}
word_counts %>%
    filter(type =='English' | type == 'Cantonese', interview=='Cantonese') %>%
    mutate(type = if_else(type=='Cantonese', 'Cantonese in\nCantonese interview', 'English in\nCantonese interview')) %>%
    group_by(word, interview, type) %>%
    summarise(n = sum(n)) %>%
    ungroup() %>%
    mutate(log_frequency = log(n)) %>%
    ggplot(aes(x=log_frequency)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_wrap(~type) +
        xlab('Log Frequency') + 
        ylab('Count')

# ggsave('../../text/figures/ch2_cantonesewordfrequency_5in.png', width = 5, height = 3)
```

English and Cantonese token frequency in the English sessions
```{r}
word_counts %>%
    filter(type =='English' | type == 'Cantonese', interview=='English') %>%
    mutate(type = if_else(type=='Cantonese', 'Cantonese in\nEnglish interview', 'English in\nEnglish interview')) %>%
    group_by(word, interview, type) %>%
    summarise(n = sum(n)) %>%
    ungroup() %>%
    mutate(log_frequency = log(n)) %>%
    ggplot(aes(x=log_frequency)) + 
        geom_histogram(binwidth = 0.5) + 
        facet_wrap(~type) +
        xlab('Log Frequency') + 
        ylab('Count')

# ggsave('../../text/figures/ch2_englishwordfrequency_5in.png', width = 5, height = 3)
```

Proportion of unknown words by talkers; an estimate of code-switching, as unknown words are almost always words in a different language.
The first plots shows the proportion based on a count, and the second based on duration.
```{r fig.width=3, fig.height=3}
unk_count <- words %>%
    mutate(is_unknown = word == '<unk>') %>%
    group_by(talker, language, is_unknown) %>%
    summarise(count_words = n()) %>%
    pivot_wider(names_from = is_unknown, values_from=count_words) %>%
    mutate(proportion_unknown = `TRUE`/(`TRUE`+`FALSE`)) %>%
    select(talker, language, proportion_unknown) %>%
    pivot_wider(names_from=language, values_from=proportion_unknown) 

unk_duration <- words %>%
    mutate(is_unknown = word == '<unk>', word_dur = word_offset-word_onset) %>%
    group_by(talker, language, is_unknown) %>%
    summarise(duration = sum(word_dur)) %>%
    pivot_wider(names_from = is_unknown, values_from=duration) %>%
    mutate(proportion_unknown = `TRUE`/(`TRUE`+`FALSE`)) %>%
    select(talker, language, proportion_unknown) %>%
    pivot_wider(names_from=language, values_from=proportion_unknown) 

ggparcoord(unk_duration, columns = 2:3,
    scale="globalminmax",
    showPoints = TRUE,
    alphaLines = 0.5,
    ) + xlab('Language') + ylab('Proportion') +
    scale_y_continuous(limits = c(0,.3),breaks = c(0,0.05,0.1,.15,.2,.25,.3))

# ggsave('../../text/figures/ch2_switchfrequency_4in.png', width = 4, height = 3)
```

## Quantity of speech

Hours of actual participant speech production -- excludes all pauses as this number is based purely on where the forced aligner identified words
```{r}
words %>%
    mutate(word_dur = word_offset - word_onset) %>%
    group_by(language) %>%
    summarise(sum(word_dur)/60/60)
```

Time actually speaking, i.e., non-silent streches summed up
```{r}
results <- dbSendQuery(
    db,
    "SELECT phone,
	       phone_onset,
	       phone_offset,
	       task,
	       task_onset,
	       task_offset,
	       phones.file
	FROM phones
	LEFT JOIN tasks ON (tasks.file = phones.file
	                    AND phone_onset BETWEEN task_onset AND task_offset)"
    )
phones <- fetch(results)

phones %>%
    filter(phone != 'sil', phone != 'sp') %>%
    mutate(dur = phone_offset - phone_onset) %>%
    group_by(file, task) %>%
    summarise(minutes = sum(dur)/60) %>%
    separate(file, into = c('talker','language')) %>%
    mutate(task = if_else(task=='sentence', 'sentences', task)) %>%
    group_by(language, task) %>%
    summarise(total_hrs = sum(minutes)/60, total_min = sum(minutes))
    #pivot_wider(names_from = task, values_from = minutes)
```

## Close SQLite database
```{r}
dbClearResult(results)
dbDisconnect(db)
rm(db, results)
```


# Participant summary

Load the data
```{r message=FALSE, warning=FALSE}
lbq_summary <- read_csv('~/Corpora/spice/info/participants/spice-lbq-summary.csv')
lbq_raw <- read_csv('~/Corpora/spice/info/participants/spice-lbq-detailed.csv')
```

```{r}
demographics <- lbq_raw %>% 
    select(!matches("language"))
```

## Places

Plot where participants have lived at different stages of their lives.
```{r fig.width=5, fig.height=4}
demographics %>%
    select(id, starts_with('Places')) %>%
    mutate(across(where(is.character), 
                  ~ str_replace_all(., 'BC |AB |ON |Quebec|CA |London |Macau|Kansai |Kanto ', ''))) %>%
    pivot_longer(-id, names_to = 'Ages') %>%
    mutate(Ages = as_factor(str_replace(Ages, 'Places lived ages', 'Ages'))) %>%
    separate(value, into = c('a','b','c'), sep = ' - ') %>%
    pivot_longer(cols = a:c, values_to = 'Place') %>%
    select(id, Place, Ages) %>%
    arrange(Place) %>%
    drop_na() %>%
    ggplot(aes(y=fct_rev(Place))) +
        geom_bar(fill='black', color='white') +
        facet_wrap(~Ages, ncol=2) +
        ylab('Country lived in') +
        xlab('Count of Participants') +
        theme(legend.position = 'none')

# ggsave('../../text/figures/ch2_placeslived_5in.png', width = 5, height = 4)
```

Plot where caretakers were primarily raised.
```{r}
demographics %>%
    select(id, contains('Caretaker')) %>%
    pivot_longer(-id, values_to = 'Place') %>%
    mutate(name = str_replace(name, 'Caretakers - Region ','')) %>%
    separate(name, into = c('Question', 'Caretaker'), sep = ' - ') %>%
    pivot_wider(names_from = 'Question', values_from = 'Place') %>%
    drop_na() %>%
    mutate(born = str_replace(born,'Manila|Ilocos ', '')) %>%
    mutate(`primarily raised` = str_replace(`primarily raised`,'Ilocos ', '')) %>%
    mutate(`primarily raised` = str_replace(`primarily raised`,'Phillippines', 'Philippines')) %>%
    mutate(`primarily raised` = str_replace(`primarily raised`,'[A-Za-z]+ China', 'China')) %>%
    mutate(`primarily raised` = str_replace(`primarily raised`,'China - Hong Kong', 'China')) %>%
    
    group_by(`primarily raised`) %>%
    summarise(n=n()) %>%
    arrange(`primarily raised`) %>%

    ggplot(aes(x=n, y=reorder(`primarily raised`, n))) +
        geom_col(fill='black', color='white') +
        xlab("Count of participants' caretakers") +
        ylab('Caretaker(s) primarily raised in')

# ggsave('../../text/figures/ch2_caretakers_3in.png', width = 3, height = 3)

```

Summarized number of caretakers who were born and/or raised in Hong Kong.
```{r}
demographics %>%
    select(id, contains('Caretaker')) %>%
    pivot_longer(-id, values_to = 'Place') %>%
    mutate(name = str_replace(name, 'Caretakers - Region ','')) %>%
    separate(name, into = c('Question', 'Caretaker'), sep = ' - ') %>%
    pivot_wider(names_from = 'Question', values_from = 'Place') %>%
    drop_na() %>%
    mutate(born = str_replace(born,'Manila|Ilocos ', '')) %>%
    mutate(born = str_replace(born,'[A-Za-z]+ China', 'China')) %>%
    mutate(bornHK = as.numeric(born=='Hong Kong'), raisedHK = as.numeric(`primarily raised`=='Hong Kong') ) %>%
    group_by(id) %>%
    summarise(bornHK = sum(bornHK)/n(), raisedHK=sum(raisedHK)/n()) %>%
    ungroup() %>%
    summarise(sum(bornHK>0), sum(bornHK==1), sum(raisedHK>0), sum(raisedHK==1))
```

## Age of learning languages

Wrangle raw LBQ for next set of plots.
```{r}
lang_nums <- lbq_raw %>% 
    select(id, matches("^Language [0-9]$")) %>%
    pivot_longer(`Language 1`:`Language 9`, names_to = "lang_num", values_to = "language") %>%
    drop_na()

tidy_languages <- lbq_raw %>% 
    select(id, matches("- Language [0-9]$")) %>%
    mutate(across(where(is.character), ~recode(.,`Excellent` = '4', 
                                               `Good` = '3',
                                               `Fair` = '2',
                                               `Elementary` = '1',
                                               `No proficiency` = NULL,
                                               `Daily` = '365',
                                               `Weekly` = '52',
                                               `Monthly` = '12',
                                               `Yearly` = '1',
                                               `Less than once a year` = '0.5',
                                               `Never` = NULL,
                                               `n/a` = NULL
                                               ))) %>%
    mutate(across(-id, as.numeric)) %>%
    pivot_longer(-id) %>%
    separate(name, into = c('question','lang_num'), sep = -10 , extra = 'merge') %>%
    drop_na() %>%
    left_join(lang_nums, by = c('id','lang_num')) %>%
    select(id, question, language, value) %>%
    mutate(question = str_replace(question, ' - $', ''),
           question = str_replace(question, '/', 'or'))

rm(lang_nums)
```

Visualize the age that each participant started learning each of their languages. This plot takes a few times to run to get acceptable overlaps in text for participants with lots of languages. 
```{r fig.height=9, fig.width=5}
tidy_languages %>%
    filter(question=='Age Started', language!='French (Quebec)') %>%
    select(Participant=id, Language=language,Age=value) %>%
    mutate(Age = if_else(Age>=20, 20, Age)) %>%
    mutate(Language = if_else(Language=='Bahasa Indonesia', 'Indonesian', Language)) %>%
    mutate(Language = if_else(Language=='Taiwanese Hokkien', 'Hokkien', Language)) %>% 
    mutate(Language = if_else(Language=='Hokkien Taiwanese', 'Hokkien', Language)) %>% 
    mutate(Language = if_else(Language=='French (France)', 'French', Language)) %>%
    arrange(Language, Age) %>%
    filter(str_starts(Participant,'VM')) %>%
    ggplot(aes(y=Age, x='', color = Age, label=Language)) +
        geom_point(size=2, shape=18) +
        geom_text_repel(max.iter=100000, min.segment.length = 0,
                        box.padding = 0.4, force=2, size=2.5, max.overlaps = 12) +
        scale_y_continuous(limits = c(-4,22), breaks = c(0,5,10,15,20), 
                           labels = c('0', '5', '10','15', '20+'))+
        scale_color_viridis_c(option='magma', begin = 0, end = 0.8) + 
        facet_wrap(~Participant, ncol = 3) +
        xlab('')+
        ylab('Age Started Learning Language (Year)') +
        theme(panel.border = element_rect(fill=NA), 
              legend.position = 'none',
              strip.text = element_text(size = 8, face = 'bold'))

# ggsave('../../text/figures/ch2_multilingualism_vm_5in.png', width = 5, height = 7)
#ggsave('../../text/figures/ch2_multilingualism_vf_5in.png', width = 5, height = 7)
```

---