%%!TEX root = dissertation.tex

\chapter{Discussion and Conclusion}\label{ch5:discussion}

What are the consequences of a shared phonetic space in the linguistic systems of bilinguals? What is shared? What is kept separate? And, how can methods couched in the study of crosslinguistic influence provide insight into these areas? Lastly, what are the implications for processing and talker identification? These questions sit at the core of this dissertation and are approached from two different angles in Chapters \ref{ch3:voice} and \ref{ch4:uniformity}, using the corpus described in Chapter \ref{ch2:corpus}. Both studies emphasize describing the structure of phonetic variation in the bilingual speech signal and work to provide a solid ground for developing hypotheses in speech perception. 

As such, this chapter proceeds as follows. Section \ref{ch5:sec:recap} recapitulates the main points of the content chapters of this dissertation, emphasizing the conclusions that are unique to each chapter. Section \ref{ch5:sec:discussion} dives into a more general discussion, highlighting how the studies conspire together to inform a broader understanding of how variation is structured in bilingual speech production. Implications for speech perception are also considered here. Section \ref{ch5:sec:limitations} makes note of limitations, and Section \ref{ch5:sec:directions} highlights some of the hypotheses that this dissertation generates but does not answer. Lastly, Section \ref{ch5:sec:conclusion} concludes by summarizing the key contributions of this dissertation to the fields of phonetics, psycholinguistics, and bilingualism.

\section{Recap}\label{ch5:sec:recap}

Chapter \ref{ch2:corpus} introduces a new speech corpus, developed as a part of this dissertation. The SpiCE corpus of Speech in Cantonese and English comprises high-quality recordings and transcripts of sentence reading, storyboard narration, and conversational interviews in each language. All 34 talkers in the corpus were early Cantonese-English bilinguals and members of the heterogeneous bilingual speech community in Vancouver, BC, Canada. Chapter \ref{ch2:corpus} documents the motivation, design, and procedures used in the creation of SpiCE. Additionally, a detailed description of the talkers and various summary statistics describing the corpus are provided. SpiCE is an open-access resource freely available to anyone interested in the data---researchers, developers, hobbyists, and the general public \citep{johnson_2021_spice}. On its own, SpiCE represents a major contribution to the study of bilingual speech production in phonetics and psycholinguistics. 

Chapter \ref{ch3:voice} describes a study on the structure of acoustic voice variation within and across languages for the talkers in the SpiCE corpus. Using a wide array of source and filter-based acoustic measurements on voiced speech in the conversational interviews, Chapter \ref{ch3:voice} investigates crosslinguistic similarity in three ways. First, the distributions of each measurement were compared across languages on a by-talker basis using Cohen's \textit{d}. The vast majority of comparisons resulted in trivial differences, indicating that talkers were, for the most part, internally consistent. Where consistent differences emerged, they mostly aligned with prior work---Cantonese tends to have lower fundamental frequency (F0) and be associated with breathier (or less creaky) voice quality than English \citep{ng_2012_ltas}. 

Second, a series of principal components analyses (PCAs) were run for each talker and language pair. In broad terms, all of the PCAs bore remarkable similarities in component structure and variance accounted for, compared to prior work \citep{lee_2019_acoustic, lee_2019_spontaneous, lee_2020_language}. This similarity across languages and talkers demonstrates a common structure among voices. The PCAs were then subjected to canonical correlation analyses to elucidate how much of the lower dimensional structure in one PCA could be accounted for by the other PCA. This analysis gets at the question of how similar two PCAs are and provides an easily interpretable metric. The results indicate that talkers are most similar to themselves across languages, compared to across-talker comparisons within or across languages. While there is some variation in the degree of similarity, the takeaway from this chapter is that voices can largely be thought of as ``auditory faces.'' Put simply, there are structural commonalities across talkers, but the greatest degree of similarity was found across languages for the same talker. 

Chapter \ref{ch4:uniformity} presents a second corpus study, focused on describing and analyzing the structure of phonetic category variation within and across languages for long-lag stops in Cantonese and English. Leveraging the uniformity framework \citep{chodroff_2017_structure}, Chapter \ref{ch4:uniformity} demonstrates that there is some structure to the relationship between voice onset time (VOT) patterns, but that the account is far less compelling than prior work on North American English \citep{chodroff_2017_structure} and L2 English varieties \citep{chodroff_2019_l2}. SpiCE talkers were wildly inconsistent with respect to the expected ordinal relationships between means for the stop categories within languages \citep{chodroff_2017_structure, cho_1999_vot, lisker_1964_vot}. That is, very few talkers produced /p/ with shorter VOT than /t/ or /k/, and likewise between /t/ and /k/. 

The second phase of the analysis considered pairwise correlations of category means for VOT within and across languages. Again, the results were far from compelling. While there were consistent moderate correlations for within-language comparisons---especially for English---the across-language correlations were weak compared to prior work \citep{chodroff_2017_structure, chodroff_2019_l2}. Their presence indicates some degree of structure but does not make for tidy conclusions. 

Chapter \ref{ch4:uniformity} ends with a Bayesian linear mixed-effects model with two primary goals. First, the model estimates the effect of language while accounting for factors known to influence VOT, such as local speaking rate and the presence of a preceding pause. Second, the model allows for evaluating the sources of variability within the model. There was a small but consistent effect of language, such that English stops are produced with slightly longer VOT than their Cantonese counterparts. The model also indicates that differences between talkers and words account for a large portion of the variation in the model, while the by-talker variability for the language and place of articulation terms, on the other hand, do not seem particularly robust. 

Two additional Bayesian models were run to investigate specific aspects of the model results. The first of these models was run without the language by place of articulation interaction. Model comparison confirms the importance of the interaction and supports an interpretation where the difference across languages is present for /t/ and /k/, but not for /p/. The second model was instead run without partial pooling for words. The result of this model is qualitatively quite different from the original model and highlights the role that lexical distributions play in shaping the results of the ordinal and correlational analyses. Words account for a lot of variation, and talkers use different distributions of words---this renders the results of the Bayesian model more reliable than the other analyses \citep[see][]{haines_2020_theoretically}. This result also highlights one of the ways that working with spontaneous speech diverges from reading. Overall, Chapter \ref{ch4:uniformity} demonstrates that there are influences of language and talker on the structure of VOT, at the very least in a coarser or ``macro'' manner. The results ultimately corroborate the ideas in the uniformity framework, albeit in a somewhat weaker manner than anticipated. 

\section{General discussion}\label{ch5:sec:discussion}

Each chapter in this dissertation includes a discussion that deals with topics central---and unique---to the study at hand. This general discussion focuses on the bigger picture and some of the implications that the two studies have for generating hypotheses in perception, as outlined in Chapter \ref{ch1:intro}. While there is substantial similarity across languages in both voice variability and the structure of long-lag stops, each study leaves room for both language and talker-indexical influences. 

\subsection{Talker-indexical and linguistic influences}

In Chapter \ref{ch3:voice}, the dual influences of talker and language show up most clearly in the canonical correlation analysis. While talkers exhibit the greatest degree of similarity when compared to themselves across languages, no talker shows perfect redundancy. That there is variability in the redundancy metric suggests influence from \textit{non}-talker-indexical sources, which could be linguistic or reflect social factors. In a strict sense, talker-indexical factors represent those that are only attributable to the talker and present regardless of circumstances or context. Linguistic factors might include cascading impacts of languages dividing up the suprasegmental space differently, and social factors might include talkers conveying different identities in each of their languages. 

Yet, while Chapter \ref{ch3:voice} does not rule out any particular type of influence, the high degree of within-talker similarity across languages emphasizes a clear role for talker-indexical components. This observation is further reflected in the comparisons of each acoustic measurement via Cohen's \textit{d}. Just over a third of the talkers in the SpiCE corpus maintain a crosslinguistic difference---in the same direction---for measures associated with F0 and non-modal voice quality. Some prior work has attempted to account for differences in F0 via the presence or absence of lexical tone. These studies, however, are not consistent with one another---some suggest lexical tone leads to lower F0 \citep{ng_2012_ltas}, and others to higher F0 \citep{lee_2017_bilingual, keating_2012_f0}. While a particular tone system certainly impacts a talker's F0 profile, there is not a clear argument for how it would impact the long-term distribution of F0 values. As such, the evidence for this argument is not yet compelling. 

If the differences in the present study were due to linguistic sources, then it would be surprising that only a third of talkers exhibited an F0 difference. A more likely account involves individual expression of identity and how that might differ across languages and ultimately manifest in F0 profiles. While speculative, there is precedent in the literature for this kind of social stratification for F0 \citep{loveday_1981_pitch, voigt_2016_between}. The lack of a clear role for linguistic influences here is further compounded by the behavior of the acoustic measurements typically associated with linguistic attributes. For example, while the patterning of F2 differences across languages largely reflects expectations around the distributions of back vowels in English and Cantonese, the direction of the effect is not consistent across talkers. While the expected difference is present for some, variability on this front and the uncontrolled nature of spontaneous speech make it challenging to draw a conclusion that reflects the group as a whole. As with the mixed-effects model results, talkers clearly differ in their lexical distributions. 

The dual influences of talker and language are perhaps more apparent in Chapter \ref{ch4:uniformity}, where individuals both maintained a VOT difference across languages and demonstrated uniformity. This outcome is not surprising, considering that Chapter \ref{ch3:voice} examined voice, while Chapter \ref{ch4:uniformity} homes in on speech sound categories. The moderate correlations for homorganic cross-language pairs indicate some level of uniformity in production---talkers with longer VOT in one language tend to also have longer VOT in the other language. Yet, at the same time, the crosslinguistic difference whereby English is characterized by longer VOT reflects a language-based influence. 

\subsection{Shared structure and consequences for perception}

Chapter \ref{ch1:intro} framed this dissertation as being concerned with the consequences of a shared phonetic space, including consideration of how the speech signal facilitates processing and bilingual talker identification. If talkers are to be consistently identified before and after a language switch, then the speech signals in each language must resemble one another to some extent. Chapters \ref{ch3:voice} and \ref{ch4:uniformity} grapple with how that resemblance plays out in production. While this dissertation does not include a perception component, the results of Chapters \ref{ch3:voice} and \ref{ch4:uniformity} have implications for how results in perception can be accounted for. 

Recall the summary of \citet{orena_2019_identifying} in Chapter \ref{ch1:intro}, in which bilingual listeners perform better at generalizing talker identification across languages compared to monolingual listeners. While all listeners in this study performed above chance, there was nonetheless a bilingual advantage. \citeauthor{orena_2019_identifying} speculated that the results could be explained via systematicity in the bilingual speech signal, whether talker-indexical, linguistic, or both. This could take the form of systematic similarity \textit{or} systematic differences across languages that bilingual listeners would have experience with. While a monolingual listener may be able to use systematic similarities, particularly in the voice, arguably, they would not be able to leverage systematic differences of any kind. Additionally, lacking familiarity with a language, these monolingual listeners may have difficulty identifying systematicity at linguistic levels, and as a result, are not able to use it. Bilinguals, on the other hand, would likely have access to all kinds of systematicity. While this speculation is useful for thinking about the results in \citet{orena_2019_identifying}, the authors do not consider production data. So, how, then, does their account fare in the context of this dissertation? Put differently, is the assumed (or proposed) systematicity present in the bilingual speech signal? 

Chapter \ref{ch3:voice} presents a strong case for the structure of acoustic voice variation and for a variety of specific acoustic measurements. Bilingual talkers show a high degree of internal structural redundancy across languages and overall similarity for the various acoustic measurements. Even in cases where a subset of talkers show a non-trivial difference across languages, few if any differences are large, especially when compared to across-talker differences. In essence, Chapter \ref{ch3:voice} supports the talker-indexical account, as similarity in the structure of voice is indeed present in the signal. It does not, however, discount the linguistic systematicity account---such information is not readily available from the results of Chapter \ref{ch3:voice}.

Chapter \ref{ch4:uniformity} offers support for both accounts. Support for the talker-indexical systematicity account arises from how much of the variation in the data arises from between-talker variability. While talkers do not necessarily show the same fine-grained patterns found in prior work \citep[e.g.,][]{chodroff_2017_structure}, there is still internal consistency. Talkers with longer VOT in one sound category tend to also have longer VOT in their other categories \citep[for a similar conclusion regarding speech rate, see][]{bradlow_2017_rate}. Support for the linguistic systematicity account arises from the mixed-effects model results. While talkers maintain a small difference across languages for VOT, this difference is not likely perceptible. And, if it were perceptible (e.g., under the right circumstances), then it would be a systematic difference that bilingual listeners are likely to have experience with. This small difference thus also supports the talker-indexical view of what listeners use to identify talkers across languages.

Returning to the broad question of what languages share in phonetic space, this dissertation provides a peek behind the curtain. Cantonese and English appear to share a lot in voice quality, despite distributional differences in the segment inventories and different roles for suprasegmental linguistic components \citep{matthews_2013_cantonese}. While such differences may be more readily apparent in shorter stretches of speech---as suggested by the passage length analysis in Section \ref{ch3:sec:passagelength}---over time, talkers appear to cover their full range. This coverage indicates that different languages make similar use of an individual's full range of acoustic voice variation and exhibit similar patterns of variation in the long run. The task of matching a new utterance up with a familiar talker, then, is one of asking whether or not the new utterance is likely to have arisen from the known range of variation for that talker.

Cantonese and English also share some aspects of phonetic category structure, albeit to a lesser extent. Chapter \ref{ch4:uniformity} demonstrates that talkers with longer VOT in one language tend to have longer VOT in the other language, even though a small distinction between the two languages was maintained. This outcome echoes results for speech rate, in which late bilinguals who are fast in their first language also tend to be fast talkers in their second language \citep{bradlow_2017_rate}. While this relationship between languages is not as simple as saying individuals use the same underlying category in each language, it does demonstrate a certain degree of shared structure.

What is clear from Chapters \ref{ch3:voice} and \ref{ch4:uniformity} is that there is considerable shared structure for listeners to use in identifying bilingual talkers. This structure provides grounding for hypotheses about what leads to a bilingual advantage in bilingual talker identification. That is, what aspect of structured phonetic variation could such an advantage actually arise from? It could stem from the variable degree of similarity for talkers across languages and bilinguals' familiarity with how voices might deviate due to social and linguistic reasons. Similarly, it could stem from bilinguals' familiarity with the variety of forms in which a particular category can be produced. While there are clear examples of this kind of sociolinguistically informed variation for initial stops in other language pairs \citep{bullock_2009_sociophonetics}, there is also evidence of metalinguistic knowledge for different sound categories in Cantonese-English bilinguals. 

A recent example juxtaposes the production of word-final stops by the talkers in the SpiCE corpus \citep{johnson_2021_language} and a lab-based study of a similar population \citep{polinsky_2018_heritage}. The corpus study demonstrates variability that skews towards Cantonese-like unreleased stops. The lab-based study, conversely, gives evidence for hypercorrection towards longer releases in English (i.e., away from the Cantonese-like behavior). By adopting the perspective of \citet{bullock_2009_sociophonetics}, this discrepancy is readily explainable via metalinguistic awareness and how bilinguals use their language in different ways when talking to their peers versus speaking in formal, monolingual, lab-based settings. The point of bringing this example up is to illustrate that bilingual listeners are not only sensitive to the fine-grained acoustics \citep{ju_2004_falling}, they are also sensitive to how form varies by communicative context. In sum, there is systematicity available in the signal for listeners---and bilingual listeners in particular---to use in tasks like talker identification. 

\section{Limitations}\label{ch5:sec:limitations}

As with any study, the results presented in this dissertation are necessarily tempered by some limitations and leave a substantial amount of variation unaccounted for. The simplest form of limitation arises from methodological decisions and is touched on in Chapters \ref{ch3:voice} and \ref{ch4:uniformity}. Both studies use corpus methods with exclusionary criteria and minimal manual inspection. In Chapter \ref{ch3:voice} this takes the form of using an automated approach to identify voiced portions of speech and a set of exclusionary criteria to discard likely errors. Chapter \ref{ch4:uniformity} relies on forced alignment, refinement via automated methods, and exclusionary criteria. Such approaches allow for the studies to be done at a larger scale but also mean that some degree of error is inevitable. The samples may include items that do not reflect the target. For example, some erroneous VOT measurements may have evaded the exclusionary criteria in Chapter \ref{ch4:uniformity}---without a rigorous manual check or manual transcription of the SpiCE corpus, the true extent of this problem will remain unknown. While it is outside the scope of this dissertation to perform such a check, prior corpus work with similar exclusionary criteria indicates that the error rate is relatively small \citep[5\%:][]{chodroff_2019_l2}.

The population studied here also presents a limitation on the extent to which the results of this dissertation can be generalized to other groups. As summarized in Section \ref{ch1:sec:bilingualism}, there is enormous variation between and among bilingual populations---the talkers in the SpiCE corpus are no exception. As described in detail in Chapter \ref{ch2:corpus}, the population studied in this dissertation represents a heterogeneous group of early Cantonese-English bilinguals. While some factors were carefully controlled for in recruitment, others were not. On the one hand, all talkers were 18-35 years old at recording, comfortable conversing in Cantonese and English, and began learning both languages well before age five (see Section \ref{ch2:subsec:participants}). On the other hand, most talkers had some knowledge of at least one additional language (e.g., Mandarin, French, etc.) and varied in their family's current or historical roots in a Cantonese-speaking homeland. 

While variability is an inherent part of the Cantonese-speaking community in Vancouver, BC, Canada---and thus justifiably included in the corpus---it does make comparisons with other bilingual communities somewhat challenging. Such comparisons are also complicated by the rather unique position of the speech community. Cantonese is reported as the ``mother tongue'' of some eight percent of Metro Vancouver census respondents \citep{statistics_2017_proportion}. As a result, Cantonese is an incredibly visible minority language in the region. As argued in \citet{chan_2020_lexically}, this population likely has more access to Cantonese than bilingual communities in other English-dominant societies \citep[e.g.,][]{bruggeman_2019_l1}. This kind of community and access to Cantonese has major implications for patterns of language use. It likely means that bilinguals have extensive experience using both Cantonese and English with a relatively wide variety of people. While there is much more to say on this topic, such detail is left for future work. Further, while this corpus could be used to explore sources of variation based on different aspects of the talkers' demographics, a lack of control on many of the potentially relevant parameters renders such approaches speculative. 

\section{Current and future directions}\label{ch5:sec:directions}

There is another kind of ``limitation'' baked into the framing of this dissertation. While the focus remains on describing and accounting for variation in speech production, the motivation arises from perception. This disjunction means that questions relating to speech perception are not answered here. Rather, the speech production results presented in this dissertation generate hypotheses for how bilingual talkers are perceived and identified. This section outlines what some of those hypotheses are. Each of the following paragraphs in this section begins with an italicized hypothesis and is followed up by its motivation. Any current research being done on the question will be noted. The first two hypotheses derive from Chapter \ref{ch3:voice} and the latter three from Chapter \ref{ch4:uniformity}.

\textit{Increased within-talker canonical redundancy across languages will facilitate crosslinguistic generalization in bilingual talker identification and discrimination. Concurrently, greater redundancy between talkers will lead to a higher chance of false alarms or confusability.} This hypothesis emerges directly from the finding that some talkers are more similar to themselves than others in the structure of their acoustic voice variation. Variability in production is thus hypothesized to be mirrored in perception. Prior perceptual work in this area often only considers a handful of talkers and uses a handful of coarse checks on the voices \citep{lavan_2019_flexible}. As a result, spurious results may be treated as being unique to the talker. This hypothesis proposes a concrete account of why talkers differ in this way. Ongoing research on bilingual talker identification and discrimination in the University of British Columbia Speech-in-Context Lab aims to address this hypothesis, along with an additional goal of better understanding the listeners' role \citep{lloy_2020_bilingual, lloy_2021_examining}.

\textit{Global shifts in voice quality dimensions across languages will not disrupt talker identification and discrimination (by bilingual listeners) when they are consistent with the crosslinguistic differences observed in the speech signal of bilingual talkers. Conversely, shifts not present in the range of voice variation patterns will be disruptive to listeners.} This hypothesis rests on a couple of different assumptions. First, that listeners are better at processing variation when they have experience with it \citep[e.g., see arguments in:][]{fricke_2016_phonetic, orena_2019_identifying}. And second, that the structure of phonetic variation presented in this dissertation is representative of that experience. This hypothesis extends the assumption that bilinguals' experience with bilingual variation will lend yet another advantage in processing variation. The rationale for proposing this hypothesis stems from how some talkers produced global differences for some of the voice measures (i.e., in F0). 

\textit{While all listeners will exhibit a processing benefit from congruent VOT within-talker across languages, bilingual listeners will be better at processing systematic differences (i.e., where VOT is consistently longer in one language compared to the other), provided they are congruent with the bilinguals' linguistic experience.} This hypothesis emerges from Chapter \ref{ch4:uniformity}, in which VOT differed slightly across languages but was also highly variable between talkers. This hypothesis ultimately echoes the proposal suggested in \citet{orena_2019_identifying}, predicting that bilinguals' knowledge of systematic differences across languages would facilitate processing. This hypothesis calls for explicit manipulation of VOT structure in a talker identification study \citep[e.g., a bilingual adaptation of][]{ganugapati_2019_structured}.

\textit{Uniformity will decrease as speech style becomes progressively less formal.} This hypothesis is supported both by prior work showing a difference in the strength of uniformity for isolated word production and reading \citep{chodroff_2017_structure}, and its comparison with the results in Chapter \ref{ch4:uniformity} which show even weaker uniformity for spontaneous speech. However, a direct comparison of spontaneous speech to either of the styles represented in \citet{chodroff_2017_structure} remains lacking, as the studies report on rather different populations and include many possible confounds. 

\textit{Full category assimilation in early bilingual speech is exceedingly rare and possibly even non-existent.} This hypothesis arises from the small difference maintained between long-lag stop categories across languages; additionally, it is supported by the arguments against compromise categories for bilinguals \citep{casillas_2021_interlingual}. This hypothesis merely extends that argument to long-lag stops that exhibit an even greater degree of crosslinguistic similarity from the outset.  

\section{Conclusion}\label{ch5:sec:conclusion}

Speech is variable, and learning to process and perceive a new talker can be characterized as learning how that talker varies. This dissertation focuses on comparing systematicity across Cantonese and English to understand how such structure might facilitate processes like multilingual talker identification. The results presented here demonstrate the presence of systematicity at two levels---acoustic voice variation and the structure of long-lag stop series. This structured variation shows evidence of both talker-indexical and linguistic influences and generates a multitude of hypotheses for future work. 

There is a balance between variation and structure at every level---talker, language, linguistic units, voice quality, and more. Working with spontaneous speech corpora is one of the best ways to gain an appreciation for this observation. It would not be possible to do this kind of research without data, and one of the largest and lasting contributions of this dissertation is the SpiCE corpus. While this dissertation just scratches the surface of what can be done with SpiCE, making the data available will help push our understanding of bilingual speech production and perception forward. 

\endinput 